Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-09-08 LaTex/paper.tex
added error band in plot, some minor review changes

In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer. The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoiled items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90 percent. Therefore, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/PdfCdf}
\caption{\label{fig:PdfCdf} Probability distribution function and cumulative distribution function of a normal distribution.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=2.5cm]{../figures/cdf_truth_uniform}
\includegraphics[width=2.5cm]{../figures/cdf_truth_broad}
\includegraphics[width=2.5cm]{../figures/cdf_truth_narrow}
\includegraphics[width=2.5cm]{../figures/cdf_truth_over}
\includegraphics[width=2.5cm]{../figures/cdf_truth_under}
\caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution:
correct prediction (``uniform''), variance overestimated (``broad''), variance underestimated (``narrow''), mean overestimated (``over''), mean underestimated (``under'')}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_example}
\caption{\label{fig:invquant_example} Inverse quantile profile plots comparing all predicted PDFs to the corresponding observed events. In the leftmost two columns, the outcome is illustrated if the estimate of the variance is biased, the center and center-right columns illustrate cases for which the mean estimation is biased, and the far right column shows the expected behavior if all predictions are correct.}
\end{center}
\end{figure}

@@

As discussed earlier, we assume that each individual sale can be described by a Poisson-like process and we assume a negative binomial distribution to model the individual probability distribution of each sales event. As a first step, we use Cyclic Boosting to predict the mean of the distribution.

@@

The second model, also based on Cyclic Boosting, is used to estimate the variance of the negative binomial distribution. The data are split into training and test set as above and in addition the mean predictions for each individual product-location-date combination are fixed in the variance model as stated by \eqn \eqref{eqn:loss_likelihood}. This effectively means that the mean predictions are created in-sample for the training period using the fully trained and validated model for the mean discussed above.

@@

\begin{figure}
\begin{center}
\includegraphics[width=4cm]{../figures/ts_item_16_store_6}
\includegraphics[width=4cm]{../figures/factors_ts}
\includegraphics[width=4cm]{../figures/factors_ts_width}
\caption{\label{fig:factor_plots} Time series of mean predictions and sales (upper left plot), constituting factors of the mean predictions (upper right plot), and constituting factors of the variance predictions (lower plot), for a specific product-location combination from beginning of February to end of April 2016. For both mean and variance predictions, the most granular constituting factors are aggregated, i.e. multiplied, for the sake of visualization.}
\end{center}
\end{figure}

@@

In many cases, scatter plots are used to study the behavior of two distributions or sets of data points visually. However, even for moderate amount of data, this approach becomes quickly difficult. To illustrate this, a sample of $(x,y)$ data points was obtained in the following way: The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$. \fig \ref{fig:scatter} shows the marginal distributions for $x$ and $y$ as well as a scatter plot of $x$ vs. $y$.

>>>>

In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer. The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoiled items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh dairy products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90 percent. Therefore, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/PdfCdf}
\caption{\label{fig:PdfCdf} Probability distribution function $f(x)$ and cumulative distribution function of a normal distribution $F(x)$.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{../figures/cdf_truth_uniform}
\includegraphics[scale=0.25]{../figures/cdf_truth_broad}
\includegraphics[scale=0.25]{../figures/cdf_truth_narrow}
\includegraphics[scale=0.25]{../figures/cdf_truth_over}
\includegraphics[scale=0.25]{../figures/cdf_truth_under}
\caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution:
correct prediction (``uniform''), variance overestimated (``broad''), variance underestimated (``narrow''), mean overestimated (``over''), mean underestimated (``under'')}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_example}
\caption{\label{fig:invquant_example} Five different inverse quantile profile plots (separated into columns by dotted vertical lines), each comparing all its predicted PDFs to the corresponding observed events. In the leftmost two columns, the outcome is illustrated if the estimate of the variance is biased, the center and center-right columns illustrate cases for which the mean estimation is biased, and the far right column shows the expected behavior if all predictions are correct.}
\end{center}
\end{figure}

@@

As discussed earlier, we assume that each individual sale can be described by a Poisson-like process and we assume a negative binomial distribution to model the individual probability distribution of each sales event. As a first step, we use Cyclic Boosting (as described in sec. \ref{sec:CB}) to predict the mean of the distribution.

@@

The second model, based on Cyclic Boosting in its negative binomial width mode described in sec. \ref{sec:cb_width}, is used to estimate the variance of the negative binomial distribution. The data are split into training and test set as above and in addition the mean predictions for each individual product-location-date combination are fixed in the variance model as stated by \eqn \eqref{eqn:loss_likelihood}. This effectively means that the mean predictions are created in-sample for the training period using the fully trained and validated model for the mean discussed above.

@@

\begin{figure}
\begin{center}
\includegraphics[width=4cm]{../figures/ts_item_16_store_6}
\includegraphics[width=4cm]{../figures/factors_ts}
\includegraphics[width=4cm]{../figures/factors_ts_width}
\caption{\label{fig:factor_plots} Time series of mean predictions (with the surrounding transparent band simplistically indicating a standard deviation as square root of the corresponding variance predictions) and sales (upper left plot), constituting factors of the mean predictions (upper right plot), and constituting factors of the variance predictions (lower plot), for a specific product-location combination from beginning of February to end of April 2016. For both mean and variance predictions, the most granular constituting factors are aggregated, i.e. multiplied, for the sake of visualization.}
\end{center}
\end{figure}

@@

In many cases, scatter plots are used to study the behavior of two distributions or sets of data points visually. However, even for moderate amount of data, this approach quickly becomes difficult. To illustrate this, a sample of $(x,y)$ data points was obtained in the following way: The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$. \fig \ref{fig:scatter} shows the marginal distributions for $x$ and $y$ as well as a scatter plot of $x$ vs. $y$.
